Make sure all equations have the same type of numbering.

Make sure page numbering is all in [] or not at all

Make sure all acronyms are in the list.
    - Pa    Pascal
    - I     Intracellular
    - E     Extracellular
    - IF    Integrate and fire
    - FHN   FitzHugh-Nagumo

Should we capitalise titles?

We should reference all figures!

Fill out empty citations!

Make sure all figures/tables are referenced as Figure (n.m)/Table (n.m). Or at least in the same way.

British or American?

Oxford comma?

---------------------------------------

The report is confusing to get into. It was very difficult to read
    - A lot of 'out of context' information
    - Explain why we are learning about cells
    - Difficult flow
    - Too much text, especially on AP (Some text could be in    the appendix(?))

The presentation made sense :)

Make it clear from the start, what we want to do

Some things are explained in a complicated way (bio part I guess), whilst some are explained in a simple way (physics)

Start by declaring that we want to couple three neurons 

Only keep information that is useful for coupling
    - Should we keep the analysis of FHN and C?

GET A RESEARCH QUESTION!!!

The text is very polished :)

KILL YOUR DARLINGS! (Could move stuff to appendix)

Make the ends meet
    - Make it fit together 
    - Red thread

We should understand everything, so we shouldn't have irrelevant things in there

There is a potential of having a very good project :)

WHAT TO DO:
    1. What is the question?
    2. How do we solve it?

We have five projects in one:
    - Comparing models
    - Coupling three neurons
    - Making a larger network
    - Seizures
    - Something else (?)

Seizures should not be a part of the core project
    - We could mention it in the introduction or in the         perspective 

CLEAN UP WORK!

Explain the terms in each equation (I feel like we are already doing this, but okay. I guess check it again).