\documentclass[printGloss]{myRUCProject}

% ========================= Preamble Packages ================================

% Margin and Notes 
\usepackage{marginnote}
\usepackage{todonotes}

\usepackage[subpreambles=true]{standalone} % Allows for independent subfile constructions

\usepackage{myTikz}

% ============================ Important Files

\import{}{customCommands}
\import{}{INP-00-glossary}
\addbibresource[glob=true]{Bibble/*.bib}

% ============================ Subfiles

% \usepackage{subfiles} % must be last usepackage

%==========================================================

\begin{comment}
    Quick ref for package commands
    
    Cleverref
    \cref[⟨options⟩]{⟨reference label⟩}
    \Cref[⟨options⟩]{⟨reference label⟩}
    The [options] is not nessecary for general usage the \cref command will detect the enviroment  that the label was made in and create the corresponding refrerence in the text
    Using \Cref produces the same result but in a  more 'official' style, like fig. 1 vs Figure 1
    
    Examples of usage;
    \cref{eq:lmao}
    \Cref{fig:penishehe}
    
    SiUnitx 
    Any general value with a unit uses the command for both the number and the unit together:
    \qty[⟨options⟩]{⟨number⟩}{⟨unit⟩}
    units are simple to input just use \⟨unit abbreviation⟩
    
    One can break up the numbers from the units by using
    \num[⟨options⟩]{⟨number⟩}
    \unit[⟨options⟩]{⟨unit⟩}
    
    Additionally, angels can be defined with
    \ang[⟨options⟩]{⟨angle⟩}
    
    Examples of usage;
    \qty{.23e7}{\candela} 
    \qty[mode = text]{1.23}{J.mol^{-1}.K^{-1}}
    \qty[per-mode = symbol]{1.99}{\per\kilogram}
    \qty[per-mode = symbol]{1.99}{\per\kg}
    \qty[per-mode = fraction]{1,345}{\coulomb\per\mole}
    

    
    Glossaries package
    \gls[<options>]{<label>} 
    \glspl[<options>]{<label>} 
    This command prints the term associated with <label> passed as its argument. If the hyperref package was loaded before glossaries it will also be hyperlinked to the entry in glossary.
    
    The pl addition prints the plural of the defined term, other than that it behaves in the same way as gls.
    \Gls[<options>]{<label>} 
    \Glspl[<options>]{<label>} 
    
    These commands are the same as the first two, just that it will print the word capitalized
    \glssymbol[<options>]{<label>} 
    

    
    General Latex commands
    for citing papers
    \cite[⟨options⟩]{⟨citation key⟩}
    for making footnotes
    \footnote[⟨options⟩]{⟨text⟩}
    
\end{comment}

%==========================================================

\title{Modeling of signal propagation through networks of neurons}
\author{ Anakin Diermayr \and  Alex Winter \and Konstantina Kalagia \and Svetlana Rusnakova \and Kenni Abild \and  Ana Helena Vieira de Melo Gomes Prata }

\begin{document}

% ======================== Title page ================================
\pagenumbering{roman}

%\maketitle
\import{./}{mainTitle.tex}
\tableofcontents

\printacronyms%

%\listoffigures \newpage


%\newpage
%\chapter{Introduction}
%\spath{Intro}


%Hello, I am a secntence \SI{56}{\um} \cite{dirac}.





\newpage
\pagenumbering{arabic}
\chapter*{Introduction} 
\addcontentsline{toc}{chapter}{Introduction}
\import{Sections/}{Intro.tex}

\part{Theory}

\import{Sections/}{background.tex}

\newpage
\chapter{Methodology}
 In order to investigate further the models in terms of applicability, scalability and accuracy, among other factors, there was a need to run computer simulations where parameters can be easily changed and their effects on models are easier to compare.
 \newline
 Python was the chosen language to do so. It is of free-access, unlike Mathlab which requires a license, more prominent than Octave(so more resources are available), and very versatile. Since, machine learning was also considered, R was deemed as not as suitable as Python. It is a statistical programming language. Moreover, the numba and numpy libraries which are based on C/C++ code make it relatively fast.\newline
 As such, python simulations were performed as this language is very well-documented, approachable and already widely used in data analysis. From the scripts, one can create phase planes and graphs to better visualize the behaviour of variables within the model.

%HAVE A SECTION WHERE WE TALK ABOUT SIMULATIONS. PYTHON-> APPROACHABLE, MUCH IN THE USE, MANY MATERIALS, WELL DOCUMENTED. 
%PRESENTATION: WE MAKE GRAPHS, TABLES, PHASE-PLANES, HOW TO MAKE SIMULATIONS(EULER METHOD, OR SOLVE THINGS). REVIEW ALEX COMMENTS!

\newpage
\chapter{Results}


\newpage
\part{Discussion}
\import{Sections/}{discussion.tex}

\newpage
\chapter{Conclusion}
%\import{Sections/}{Conclusion.tex}


\appendix
\part*{Appendix}
\addcontentsline{toc}{part}{Appendix}

\chapter{Images}\label{ap:sfd}
\import{Sections/}{Appendix}

\end{document}
