\documentclass[../../Orator.tex]{subfiles}

\begin{document}

The attempt to find an equation which can accurately describe the behavior of a chosen system is one of the most fundamental aspects of mathematics.

A model is a series of mathematical equations capable of replicating the behavior of a system. A role of these models, known as dynamic systems, is to bring light in the gaps of understanding and explain the underlying mechanisms behind some function. 
The construction of a model requires equations complex enough to accurately describe the dynamics of interest yet, preferably, simple enough so that mathematical tools exist to analyze the equations\footnotemark. 
\footnotetext{The best model of a cat is a cat. Preferably the same cat\footnotemark.-`\textit{Philosophy of Science, 1945, Arturo Rosenblueth \tbr{1900-1970}}'}
There exists any number of limitations in the pursuit of this task, some inherent to the system, others inherent to our modern construction of mathematics.
\footnotetext{If man could be crossed with the cat, it would improve man, but it would deteriorate the cat.-`\textit{Mark Twain}'}

\section{Dynamic System Fundamentals}
When looking at the brain in abstract form it can be seen as a multi-dimensional \gls{gls:dynamical system} governed by an independent set of system variables, such as neuronal membrane potentials, which change in time based on a set of \gls{gls:deterministic} equations with system parameters that either do not change in time (e.g. maximal conductance of the ion
channels on the neuronal membrane) or whose evolution happens on a much slower time scale relative to the evolution of the system variables ~\cite{STEFANESCU2012748}.


Fundamentally, a \gls{gls:dynamical system} is a system that evolves over time. This system is dependent on fixed rules which govern the evolution of the system. Thus a \gls{gls:dynamical system} is the opposite of a static system, which, unlike \gls{gls:dynamical system}s, does not depend on the past inputs. A simple example of a \gls{gls:dynamical system} can be that of a pendulum swinging back and forth \cite{}. 

There are different types of \gls{gls:dynamical system}s. A system can be \gls{gls:deterministic} or \gls{gls:stochastic}, \gls{gls:discrete} or \gls{gls:continuous}, \gls{gls:linear} or \gls{gls:nonlinear}, and \gls{gls:autonomous} or \gls{gls:non-autonomous} \cite{}. 

If a system is \gls{gls:deterministic}, it is possible to predict each following state given an initial state. It is also possible for a system to have an element of randomness to it. In this case, the system is said to be \gls{gls:stochastic} \cite{}. 

A \gls{gls:discrete} system is one where there is only measured a position of the integer values of time. On the other hand, is a \gls{gls:continuous} system, where the positions are measured \gls{gls:continuous}ly (ie. for every possible time) \cite{}.

A \gls{gls:linear} system is one where it is only composed of \gls{gls:linear} functions and a \gls{gls:nonlinear} system contains at least one \gls{gls:nonlinear} component \cite{}.

Lastly, there is the difference between an \gls{gls:autonomous} system, which is one where it does not depend on the independent variable \(t\) \cite{}, and a \gls{gls:non-autonomous} system, which does depend on \(t\) \cite{}.

%The 
%Some of these parameters

\begin{comment}
    \begin{split}\left[\begin{array}{ccll}
    {\displaystyle \frac{du}{dt}} &=& u\left(1-u^{2}\right)-w+I \equiv F(u,w)\\[.2cm]
    {\displaystyle \frac{dw}{dt}} &=& \varepsilon \left(u -0.5w+1\right) \equiv \varepsilon G(u,w)\, ,\\
    \end{array}\right.\end{split}
    [ref:https://neuronaldynamics-exercises.readthedocs.io/en/latest/exercises/brunel-network.html]
\end{comment}

When analysing differential equations \textit{phase space} analysis is an important concept and tool. Phase space is the space spanned by the system's variables and the trajectories it contain are those variables traversing time. When a neuron transitions from rest to firing a qualitative change in the dynamical behavior of the system will occur. A qualitative change can be associated with phenomena such as \textit{bifurcations}, \textit{phase transitions}, emergence of (new) \textit{attractors}, or shifts in stability. 

Bifurcations constitutes critical points where the stability of a system changes, leading to abrupt transitions in the system's behavior. As will be shown these transitions can provide insights into the onset of seizures and their characteristics [link to network analysis section]






Notes
- Time delays can be the source of instabilities and bifurcations in \gls{gls:dynamical system}s and are frequently observed in biological systems such as neural networks.
 to understand how system behavior changes as a parameter is varied
with qualitative change in the dynamical behavior of the system we mean 
- The bifurcation constitutes a 'dividing event' and is associated with modifications to the system parameters such as membrane capacitance and ion channel parameters etc. ~\cite{STEFANESCU2012748}.
- Limit cycle


Bifurcations are critical points where the stability of a system changes, leading to abrupt transitions in the system's behavior. These transitions can provide insights into the onset of seizures and their characteristics.
Identifying bifurcation points in neural systems helps to understand the mechanisms underlying the transition from normal brain activity to a seizure state.

\subsubsection{Dimensionality reduction}
A neuron model, such as the \gls{hh} model, uses a system of four differential equations
to simulate a complex network this can quickly amount to ...



\subsection*{Modeling Neuronal Networks}
A complex system can be described by a network or a graph with complex topology, whose nodes are the elements of the system and whose edges represent the interactions among them. `One significant recent discovery in the field of complex networks is the observation that a number of large-scale and complex networks are scale-free, that is, their connectivity distributions have the
power-law form'~\cite{wang2002synchronization}.


...



When analysing seizures, bifurcation points helps to understand the mechanisms underlying the transition from normal brain activity to a seizure state. The change in parameter can be factors like synaptic strength, excitability of neurons, or network connectivity 
~\cite{gerstner2014neuronal}. 





    
\subsubsection{Network connectivity and topology}


Modeling brain processes can be advantages because of the difficulties  performing certain experiments and it is therefore possible to simulate certain scenarios without the need of conducting large scale exploratory experiments. Aspects like ethical barriers...

The practical barriers to simulating brain processes are to do with the modeling framework of the particular question of interest together with the computational resources needed to run complex simulations. Fortunately Moore's Law are still holding and with the emergence of AI the modeling framework have been propelled foreword. Large-scale brain simulation projects, like the Human Brain Project\footnote{\url{https://www.humanbrainproject.eu/en/brain-simulation/}}, leverage AI and high-performance computing to simulate the behavior of billions of neurons and their connections.


\subsection*{Seizure}
Seizures can be primarily interpreted as a dynamical disease \cite{da2003epilepsies, milton2010epilepsy} and computational models have been successful in gaining insight into and generate hypothesis to the cellular and network level brain mechanisms of seizures \cite{bazhenov2008cellular}.

\subsubsection*{Definition and classification of seizures}



\subsubsection{The role of neuronal networks in seizure generation}
(or network effects )

- network properties influence seizure dynamics --> connectivity and coupling strengths


Some notes:
- The role of neuronal networks in seizure generation
- Models for simulating seizures
- A chaotic process can be classified according to its fractal dimensions and Lyapunov exponent.
- Identifying bifurcation points in neural systems 



When modeling neuronal interactions, and in particular modeling of seizures, the model must reflect the fact that thousands of neurons need to interact in order to display seizure-like activity. 
The first major roadblock becomes the quantity of neurons involved, it is not possible to measure the activity of every individual neuron, even assuming one could do so, it would be a monumental task to extract useful information from the bulk. 
Therefore we must be smart about creating a model, choosing the `right' approach to simplify the equations can save astronomical amounts of time and effort.
\begin{enumerate}
    \item \textbf{Model average activity}
    \item \textbf{Reduce parameter space}
    \item \textbf{Tackle a smaller problem}
\end{enumerate}


A system can be modeled \gls{gls:deterministic}ly or contain \gls{gls:stochastic}ly. When collecting data, measurements will rarely be completely \gls{gls:deterministic}. This is not because forces act in unpredictable manner, rather that there exists far too many elements in the `real' world for any model to fully account for them. 
As a consequence, many models will incorporate both \gls{gls:deterministic} and \gls{gls:stochastic} elements to more accurately reflect the observed data.
\begin{enumerate}
    \item \textbf{Model simplification}
    \item \textbf{\Gls{gls:stochastic} activity}
    \item \textbf{\gls{gls:stochastic} Inputs}
\end{enumerate}



\newpage
\section{Physiological Parameters}
It's important to determine what is even being modeled, which features can be parameterized and which can not.

\subfile{parameters}


\newpage
\section{Building up to Hodgkin-Huxley}

The \gls{hh} Model of Neuron Action Potential is regarded as being a part of the great achievements of 20th-century biophysics. Receiving the Nobel Prize in Physiology or Medicine for what is inarguably an incredible feat of human ingenuity. However, the authors of this paper are students, and therefore such a sophisticated model might be beyond the direct reach of abilities within the given timeframe. 
As such, the exploration begins by investigating simplified \gls{hh} derived models to see what insights can be built.

\subsection{FitzHugh-Nagumo}
``The models in this category are highly simplified toy models that qualitatively describe the membrane voltage as a function of input. They are mainly used for didactic reasons in teaching but are not considered valid neuron models for large-scale simulations or data fitting''.\footnote{Wikipedia} 
%\subsubsection{Didactic toy model}
\begin{subequations}\label{eq:FitNagSys}
    \begin{align}
        \ode{V}{t} &= V - \frac{V^3}{3} - w + I_{ext} \\ 
        \tau\ode{w}{t} &= V - a - b\,w 
    \end{align}
\end{subequations}


\subsection{Chay}
In 1985, T.R. Chay proposed a model of three-dimensional \gls{gls:nonlinear} differential equations based on the \gls{hh} model to study chaotic behavior and show ionic events in excitable membranes. 
\begin{subequations}\label{eq:ChaySys}
    \begin{align}
        \ode{V}{t} &= g_\rmm{I}  m^3_\infty h_\infty \br{V_\rmm{I} - V} + g_\rmm{K, V} n^4 \br{V_\rmm{K} - V} + g_\rmm{K, C}  \frac{C}{1+C}\br{V_\rmm{K} - V} + g_\rmm{L} \br{V_m - V_\rmm{L}} \\ 
        \ode{n}{t} &= \frac{n_\infty - n}{\tau_n} \\
        \ode{C}{t} &= \rho \, \sbr{m^3_\infty h_\infty \br{V_c - V} - k_C C}
    \end{align}
\end{subequations}

where \textit{V}, \textit{n}, and \textit{C} are membrane potential, probability of the voltage-sensitive \(\Kp\) channel, and intracellular concentration of \(\Capp\) ions, respectively. The Chay model parameters are adopted from \textit{`paper'} and collected in Table

The \(m_\infty\), \(h_\infty\), and \(n_\infty\) are calculated by \(y_\infty = \alpha_y / \br{\alpha_y + \beta_y} \) formula, and the explicit expressions for 
\(\alpha_m, \beta_m, \alpha_h, \beta_h, \alpha_n, \beta_n\), and \(\tau_n\) are given by:
\begin{align*}
    \alpha_m &= 0.1 \frac{ 25 + V }{1 - \exp{-0.1 \, V - 2}}, &
    \alpha_h &=  0.07 \exp{-0.05\,V -2.5}, &
    \alpha_n &= 0.01 \, \frac{ 20 + V }{ 1 + \exp{-0.1 \, V - 2}} \\
    \beta_m  &= 4 \exp{-\br{\frac{ V + 50 }{ 18 } } }, &
    \beta_h  &= \frac{ 1 }{ 1 + \exp{-0.1 \, V - 2}}, &
    \beta_n  &= 0.125 \exp{- \frac{V + 30}{80}}, \\
    \tau_n &= \frac{1}{ r_n \, \br{\alpha_n + \beta_n} }
\end{align*}



\newpage
\subsection{Hodgkin-Huxley}\label{sec:HHMeth}

The \gls{hh} model of Action Potential is a system of \gls{gls:nonlinear} differentiable equations with four state variables with respect to time, \(V_m\br{t}, \ n\br{t}, \ m\br{t}, \ h\br{t}\)~\cite{HodHux1952}. The model is built from approximating the characteristics of excitable cells, such as neurons, to a circuit-like construct \cref{fig:MembraneCircut}. 
Mathematically, the flow of current is represented as:
    \begin{align}
        I_c &= C_m \ode{V_m}{t} \\
        \intertext{and the current through a given ion channel is the product of that channel's conductance and the reversal potential for the specific ion}
        I_\bfm{ion} &= {g_\bfm{ion}}(V_\rmm{m}-V_i)  
    \end{align}
        where \(V_\bfm{ion}\) is the reversal potential of the specific ion channel. Thus, for a cell with sodium and potassium channels, the total current through the membrane is given by:
    \begin{subequations}\label{eq:HodHuxSys}
        \begin{equation}
            I = C_m \ode{V_m}{t} + \bar{g}_\rmm{K} n^4 \br{V_m - V_\rmm{K}} + \bar{g}_\rmm{Na} m^3 h \br{V_m - V_\rmm{Na}}  + \bar{g}_\rmm{L} \br{V_m - V_\rmm{L}} 
        \end{equation}
The duo of Hodgkin and Huxley divined the model from observing the smooth current changes in terms of pores or channels that were either open or closed, and by using a statistical approach generated predictions for the probability of channels being open. 

Accounting for the current through a any given ion channel is the product of the channel's conductance and equilibrium potential for the specific ion.
These `gating' variables \(m\) and \(h\), or \(n\) describe the time dependent kinetics of the voltage
        \begin{align}
            \ode{n}{t} &= \alpha_n \br{V_m} \br{1-n} - \beta_n \br{V_m} n \\
            \ode{m}{t} &= \alpha_m \br{V_m} \br{1-m} - \beta_m \br{V_m} m \\
            \ode{h}{t} &= \alpha_h \br{V_m} \br{1-h} - \beta_h \br{V_m} h 
        \end{align}
    \end{subequations}

The ion channel activation/inactivation\footnotemark~probabilities, denoted by \(\alpha_p, \, \beta_p : \, \br{n,m,h} \in p\), are defined such that:
\begin{align}
    \alpha_p\br{V_m} &= p_\infty \br{V_m} / \tau_p \\
    \beta_p\br{V_m}  &= \br{1 - p_\infty \br{V_m}} / \tau_p 
\end{align}
With \(p_\infty\) and its inverse \(1-p_\infty\) being the steady state values for activation and inactivation respectively~\cite{}. 
\footnotetext{\underline{a}ctivation gives \(\alpha\), \underline{b}nactivation gives \(\beta\), really these are the obvious choices}


In the original paper by Hodgkin and Huxley, the relationships of \(\alpha_p, \text{ and }\, \beta_p\) were defined as:

\begin{minipage}{.45\textwidth}
\begin{equation*}
n \implies 
\begin{cases}
    \alpha_n \br{V_m} &= 0.01 \, \cfrac{ 10 - V }{ \exp{ 10 - V } - 1} \\
    \beta_n \br{V_m}  &= 0.125 \, \exp{-\cfrac{V}{80}} 
\end{cases}
\end{equation*}
\end{minipage}
~
\begin{minipage}{.45\textwidth}
\begin{equation*}
m \implies 
\begin{cases}
    \alpha_m \br{V_m} &= 0.1 \cfrac{ 25 - V }{\exp{\cfrac{25-V}{10}} - 1} \\
    \beta_m \br{V_m}  &= 4 \, \exp{-\cfrac{V}{18}} 
\end{cases}
\end{equation*}
\end{minipage}

\begin{minipage}{.45\textwidth}
\begin{equation*}
h \implies 
\begin{cases}
    \alpha_h \br{V_m} &=  0.07 \, \exp{-\cfrac{V}{20}} \\
    \beta_h \br{V_m}  &= { \cfrac{1}{\exp{\cfrac{30 - V}{10}} + 1}}
\end{cases}
\end{equation*}
\end{minipage}
~
\begin{minipage}{.45\textwidth}
    \hfill
\end{minipage}

Where \(V = V_\rmm{rest} - V_m\) represents the polarization in \unit{\milli\volt}

In order to characterize the ion-channels, the equations can be fitted to voltage clamp\footnotemark~data.

\footnotetext{An assay that measures the flow of current through a neuronal cell membrane by `clamping' the potential at an unchanging value.}

\begin{comment}
    Breaking down the equation we find the following regions:
    \begin{equation*}
        I = 
        \overbrace{C_m \ode{V_m}{t}}^{\mathclap{\text{rate of change of voltage scaled by capacitance}}} + 
        \bar{g}_\rmm{K} n^4 \br{V_m - V_\rmm{K}} + 
        \bar{g}_\rmm{Na} m^3 h \br{V_m - V_\rmm{Na}}  + 
        \bar{g}_\rmm{L} \br{V_m - V_\rmm{L}} 
    \end{equation*}




    \begin{figure}[h]
        \centering
        \begin{tikzpicture}
            
            \draw (0,0) circle (0.25);
            
        \end{tikzpicture}
        \caption{fart}\label{<label>}
    \end{figure}
\end{comment}


\end{document}